---
title: "Statistical Methods 6 Protfolio"
author: "Yousef Aldawoud"
date: "10th June 2020"
output:
  html_document:
    theme: cosmo
    highlight: tango
---


**StudentID : 18038023**

[Web version](https://yousef-aldawoud.github.com/statistics-portfolio)

[GitHub](https://github.com/yousef-aldawoud/statistics-portfolio)

# Analyzing score data set of a color game

## Background

Playing video games was one of the hobbies I grew up doing and loving. Playing video games can give you some enjoyment while challenging your skills. In this analysis I try to answer some questions that might show some behavior difference between gamers and non-gamers 

## Introduction

Color game is a game where you have to pick the right color within the time limit.
There are two modes of the game.

**Normal mode**

The game will ask you to choose a color from 4 options within 3 seconds. The default number of questions is 20. It can be changed but in this case we stuck to 20 question per person

**Example :** <span style='color:white;background:red;padding:2px;'>red</span>

**Shuffled colors mode**
The game will ask you to choose a word from 4 options where colors and words are shuffled .

**Example :** <span style='color:white;background:green;padding:2px;'>red</span>

The default number of questions is 20. It can be changed but in this case we stuck to 20 question per person.

You can play the game on this [link](https://yousef-aldawoud.github.io/color-game/)

### Dataset 

The data set collected shows the scores of the games played twice in both modes.
It also show you the age, gender, and if the person is a gamer (plays video-games casually)
The dataset is assigned IDs to keep the names anonymous.

**Reading dataset**

```{r data_read, echo=FALSE}
library(knitr)
game_data = read.csv('data.csv');
kable(head(game_data));
```

#### Attributes

**ID**: a sudo id for the users to keep them anonymous.

**First time** : How much the player scored in the first try (normal mode).

**Second time** : How much the player scored in the second try (normal mode).

**First time with shuffle** : How much the player scored in the second try (shuffled mode).

**Second time with shuffle** : How much the player scored in the second try (shuffled mode).

**Gamer** : plays video games casually.

**Age** : player's age.

**Gender** : player's gender (m for male, f for female).





#### Data summary

```{r data_summary, echo=TRUE}
kable(summary(game_data));
```

The observations are as follows:

There are `17` observations. 
9 males 8 females and 9 gamers (people who casually play games) and 8 people who don't play games.


### First score in the game:

In the data set we have many people who scored high scores from the start such as 20 and 19. Given it is their first time playing the game
Which made me question weather all players are going to score the more in the second try than the first try

My null hypothises is that the number of scores in the second try is less than the first try

```{r t_test, echo=TRUE}
observed_first_score <-game_data$first.time;
observed_second_score <-game_data$second.time;
boxplot(observed_first_score,observed_second_score, names = c('first time','second time'));
pValue = t.test(observed_first_score,observed_second_score, alternative = "less")$p.value;
cat("p-value = ",pValue)
```

We fail to reject the null hypothises because the p-value is more than 0.05


### Casual gamers 

One of the questions I asked people before starting the game is weather they usually play video games as a hobby. 
Video games being one of the hobbies that I enjoy I always hear that there are more male players than female players.
Which I wanted to see weather this is a true statement or just a false stereotype.
My null hypothises is that there being a casual gamer is dependant from the gender variable.

```{r fisher_test,echo=TRUE}
contingencyTable = table(game_data$gender,game_data$Gamer);
kable(contingencyTable);
fisher.test(contingencyTable)
```

since the p-value is less that 0.05 we can reject the null hypothises. It seems that that most gamers are males.